---
title: "K Means Clustering"
output: word_document
---
  
```{r, include = FALSE}
library(tidyverse)
library(tidymodels)

#library(cluster) #algorithms for clustering
#library(factoextra) #visualization
```


```{r}
trucks=read_csv('trucks.csv')

```


Relationship between Distance and Speeding
```{r}
ggplot(trucks, aes(x=Speeding,y= Distance)) + geom_point() 

```

We see that above scatter plot shows at least two natual clusters.



```{r}
kmeans_recipe = recipe(~ Distance + Speeding, trucks) 
trucks_dummy = kmeans_recipe  %>%
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy,trucks) #prepares the recipe

trucks_cleaned = bake(trucks_dummy,trucks) 
summary(trucks_cleaned)

```

```{r}
set.seed(64)
clusts = 
  tibble(k = 2) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Speeding, y = Distance)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

We see that two clusters are at a very distant from each other however have too much spread within and hence we should look for more number of clusters.


```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))

p1 = 
  ggplot(assignments, aes(x = Speeding, y = Distance)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

From above plot, k = 4 looks more appropriate ensuring clear distinction between clusters.



```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point()
```


In above graph, we see that SS reduction is insignificant after k =4 hence that would be approriate value.



```{r}
set.seed(64)
clusts = 
  tibble(k = 4) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))

p1 = 
  ggplot(assignments, aes(x = Speeding, y = Distance)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

Now we see 4 different clusters which looks to have centroid at a significance distance from each other and at least 3 clusters are moderately or highly densed ensuring lower within SS.